---
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
---

```{r}
library(tinytex)
library(tidyverse)
library(dplyr)
dfA <- ELEC_copy
dfB <- ELEC2
```

##MILESTONE2.A MIILESTONE2.A EDA journey First we analyse the First Dataset, % of Elctricity per country per year. -Which Data set columns can be clean in order to do a more concise analysis? -

```{r}
summary(dfA)
```

With a brief summary we noticed the following:

-There is no data for the years: 1960-1981 and 2021 -First Register is 1990 with 158 missing values -Last register in 2020 has only 2-3 missing values , we see this since 2009

We also can get the following insights:

We Clean the Data in order to eliminate redundant columns:

```{r}
dfa1 <- select(dfA,-('1960':'1989'),-'2021',-'Indicator Name',-'Indicator Code')
```

First thing to analyse is to compare the data of 1990 (Our starting point) and 2020 (ending point)

Which are the overall differences between 1990 and 2020? What is the country with less % of population with electricity in 1990? What is the country with less % of population with electricity in 2020?

```{r}
summary(dfa1[c('1990','2020')])
min2020 <- which.min(dfa1$'2020')
CNM2020 <- dfa1$'Country Name'[min2020]
min1990 <- which.min(dfa1$'1990')
CNM1990 <- dfa1$'Country Name'[min1990]
#Which was the least covered country in 1990?
CNM1990
#Which was the least covered country in 2020?
CNM2020
```

South Sudan numbers are quite low, we want to do a deep dive in order to understand it better:

```{r}
dfa1 %>% filter(dfa1$'Country Name' == 'South Sudan')
```

We can see South Sudan has data since 2006 were it started with 0,6%, and they been improving slowly. The value is surprisingly low, but it´s consistent trough the years and it´s improving.

We also want to Deep dive in Zambia, have they improved?

```{r}
dfa1 %>% filter(dfa1$'Country Name' == 'Zambia')
```

We see Zambia has improved from 13.90% in 1990 to 44.5% in 2020. Despite the 2020 number is low, they have improved over the years.

We are from Chile and Colombia, we are interested to see our nation's statistics:

```{r}
dfa1 %>% filter(dfa1$`Country Name` %in% c("Chile","Colombia"))
```

We see a positive trend, with 100% coverage in 2020.

How many countries in 2022 are below the global average of 2022 (86.412%)? What is the % of those countries?

```{r}

avg2020 <- mean(dfa1$`2020`,na.rm = TRUE)
dfa1 <- dfa1 %>%
mutate(
global_average=
  ifelse(
    dfa1$`2020` < avg2020
    ,1,0))
sum(dfa1$`global_average`,na.rm=TRUE) 
sum(dfa1$`global_average`,na.rm=TRUE)/length(dfa1$`global_average`) * 100


```

From those nations below the global average(2020), what is the average accesses of electricity coverage?

```{r}

dfa1_below <- dfa1 %>% filter(dfa1$`global_average`== 1)
mean(dfa1_below$`2020`,na.rm = TRUE)
```

Which are the top 10 nations with lowest % of electricity coverage in 2020? What is the difference between 2000 and 2020?

```{r}

dfa1 <-dfa1 %>% 
  mutate(
    difference_between_2020_2000 = dfa1$"2020" - dfa1$"2000"
  )
dfbottom = dfa1[order(dfa1$"2020", decreasing = FALSE), ] %>% 
  select("Country Name", "2020","difference_between_2020_2000")
dfbottom %>% head(10)
```

Are there any nations which had decrease the percentage decrease over that period? Which are the countries?

```{r}

dfa1 <-dfa1 %>% 
  mutate(
    difference_between_2020_2000 = dfa1$"2020" - dfa1$"2000"
  )
dfa1 %>% 
  arrange(difference_between_2020_2000) %>% 
  select(`Country Name`, difference_between_2020_2000) %>% 
  head(3)

```

How many countries are over 95% of coverage? What is their % ?

```{r}
dftop <- dfa1 %>%
  filter(`2020` > 95) %>% 
  nrow()
print(dftop)
sum(dftop)/length(dfa1$"2020") *100
```

It´s impressive to see 65% of the nations are over 95%.

##MILESTONE2-B: In order to build interesting tables, We want to understand the second database in our database:

```{r}
summary(dfB)
unique(dfB$Region)
unique(dfB$IncomeGroup)
```

We see we have 8 Unique region and 4 income level. We want to understand the data by Region and Income Level so we joined de data frames

```{r}
#We will leftjoin the dataset in order to understand the data by nation at a region and income level
dfull <- dfa1 %>% 
  left_join(dfB, 
            by = 'Country Code') 
```

For the first table we want to see a summary of the general information by Region.

```{r}
library("kableExtra")
library("knitr")
RegionGroup <- dfull %>% 
group_by(Region) %>% 
  summarise(
    Total_Nations = length(`2020`),
    Mean_2020 = mean(`2020`,na.rm = TRUE),
    Mean_2010 = mean(`2010`,na.rm = TRUE),
    Mean_2000 = mean(`2000`,na.rm = TRUE),
    Below_Global_AVG = mean(`global_average`,na.rm = TRUE))%>%
  arrange(desc(Mean_2020),.by_group = FALSE)%>%
  filter(Region != "")
kable(RegionGroup,caption= "<ul>
                  Summary Statistics by Region:
                  <li>In 2020 they only region below 90% is Sub-Saharan Africa</li>
                  <li>All regions have improved trhough the years</li>
                  <li>Almost 90% of Sub-Saharan Africa are below global avg</li>") %>%
  kable_styling()
```

For the second table we want to see a summary of the general information by Income level.

```{r}
IncomeGroup <- dfull %>% 
group_by(IncomeGroup) %>% 
  summarise(
    Total_Nations = length(`2020`),
    Mean_2020 = mean(`2020`,na.rm = TRUE),
    Mean_2010 = mean(`2010`,na.rm = TRUE),
    Mean_2000 = mean(`2000`,na.rm = TRUE),
    Below_Global_AVG = mean(`global_average`,na.rm = TRUE))%>%
  arrange(desc(Mean_2020),.by_group = FALSE)%>%
  filter(IncomeGroup != "")
kable(IncomeGroup,caption= "<ul>
                  Summary Statistics by Income Group:
                  <li>In 2020 they Lower income levels are below 85%</li>
                  <li>All regions have improved trhough the years</li>
                  <li>More than 90% of Low Income level are below global avg</li>") %>%
  kable_styling()
```

We want to do a third table in order tO understand the situation of Sub-Saharan Africa per Income level

```{r}
RegionGroupIncome <- dfull %>% 
group_by(Region,IncomeGroup) %>% 
  summarise(
    Total_Nations = length(`2020`),
    Mean_2020 = mean(`2020`,na.rm = TRUE),
    Mean_2010 = mean(`2010`,na.rm = TRUE),
    Mean_2000 = mean(`2000`,na.rm = TRUE),
    Below_Global_AVG = mean(`global_average`,na.rm = TRUE))%>%
  arrange(desc(Mean_2020),.by_group = FALSE)%>%
  filter(
    (Region != "") & 
    (IncomeGroup != "") &
    (Region == "Sub-Saharan Africa")
    )
kable(RegionGroupIncome,caption= "<ul>
                  Summary Statistics by Income Group in Sub-Saharan Africa :
                  <li>Critical Situation for Lowe Income</li>
                  <li>100% of lowest income are below Global Average</li>
                  ") %>%
  kable_styling()
```

##Milestone2C

```{r}
RegionGroup <- RegionGroup %>% arrange(desc(Mean_2020))
ggplot(RegionGroup, aes(reorder((x = Region),Mean_2020), y = Mean_2020,fill = Region == "Sub-Saharan Africa", width = 0.8)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("green", "red"), 
                    breaks = c(FALSE, TRUE),
                    labels = c("Other regions", "Sub-Saharan Africa")) +
  labs(title = "% of Population with Access to electricity in 2020 per Region",
        x = "Region", y = "% Electricity", caption = "Sub-Saharan region is the only one facing electricity challenges in 2020") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         plot.margin = unit(c(0.1,0.5,0.1,0,1,0.1), "cm"),
         plot.caption = element_text(hjust = 0, margin = margin(10, 0, 10, 0))) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20), labels = paste0(seq(0, 100, by = 20), "%")) +
  guides(fill = guide_legend(override.aes = list(width = 0.8, size = 2)))
```

```{r}
RegionGroup <- RegionGroup %>% arrange(desc(Mean_2020))
ggplot(IncomeGroup, aes(reorder((x = IncomeGroup),Mean_2020), y = Mean_2020,fill = IncomeGroup == "Low income", width = 0.8)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("green", "red"), 
                    breaks = c(FALSE, TRUE),
                    labels = c("Other Incomes", "Low income")) +
  labs(title = "% of Population with Access to electricity in 2020 per Income Level",
        x = "Region", y = "% Electricity", caption = "Low income level is the only one facing electricity challenges in 2020") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         plot.margin = unit(c(0.1,0.5,0.1,0,1,0.1), "cm"),
         plot.caption = element_text(hjust = 0, margin = margin(10, 0, 10, 0))) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20), labels = paste0(seq(0, 100, by = 20), "%")) +
  guides(fill = guide_legend(override.aes = list(width = 0.8, size = 2)))


```

```{r}
ggplot(RegionGroupIncome, aes(reorder((x = IncomeGroup),Mean_2020), y = Mean_2020,width = 0.8)) + 
  geom_bar(stat = "identity")  +
  labs(title = "% Access to electricity in 2020 per Income Level from Sub-Saharan Africa ",
        x = "Region", y = "% Electricity", caption = "There is clear realtion with income and Low Income is in a critical situation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         plot.margin = unit(c(0.1,0.5,0.1,0,1,0.1), "cm"),
         plot.caption = element_text(hjust = 0, margin = margin(10, 0, 10, 0))) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20), labels = paste0(seq(0, 100, by = 20), "%")) +
  guides(fill = guide_legend(override.aes = list(width = 0.8, size = 2)))
```

```{r}
dfa1_subset <- dfa1[,c("Country Name", as.character(1990:2020))]

dfa1_long <- dfa1_subset %>%
  pivot_longer(cols = -c("Country Name"), names_to = "Year", values_to = "Percentage")

dfa1_long$Year <- as.numeric(dfa1_long$Year)

dfa1_long <- dfa1_long %>% na.omit();
ggplot(dfa1_long[dfa1_long$`Country Name` == "Sub-Saharan Africa",], 
       aes(
         x = Year, 
         y = Percentage,
         color = "Sub-Saharan Africa",
         group = 1)) +
geom_line(linewidth = 1.3) +
geom_line(data = dfa1_long[dfa1_long$`Country Name` == "World",], 
          aes(x = Year, 
              y = Percentage,
              color = "World",
              group = 1), 
          linewidth = 1.3) +
ggtitle("Access to Electricity in Sub-Saharan Africa") +
xlab("Year") +
ylab("Percentage of the Population") +
scale_y_continuous(limits = c(0, 100)) +

labs(subtitle = "Despite the situation is critical, we see an evoulution trhough time") +
guides(color = guide_legend(title = "Country"))
```
